Parental Migration and Left-Behind Children's Cognitive and Academic Performances

### 2022-06-21 MScThesisText.Rmd
write up main text, set matching criteria, add summary tables

---
title             : "Parental Migration and Left-Behind Children's Cognitive and Academic Performances"
shorttitle        : "Left-Behind Children"

author: 
  - name          : "Jiancheng Gu"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "De Boelelaan 1105, 1081 HV - Amsterdam, The Netherlands"
    email         : "j4.gu@student.vu.nl"

@@

floatsintext      : no
linenumbers       : no
draft             : no
mask              : no

@@

```{r setup, include = FALSE}
load('.Rdata')
library("papaja")
library("modelsummary")
r_refs("references.bib")
```

@@

Millions of children are resided in their home communities but separated from their parents as either one or both parents have migrated for work [@antia2020]. These are the so-called "left behind children" who are common in migrant-sending regions in the Global South. The global size of left behind children population is unavailable, but in-country surveys provide some estimation. [@dewaard2018] analyzed the Mexican and Latin American Migration Projects and estimated that around 15 percent of children were left behind in Mexico, El Salvador, Nicaragua, and Puerto Rico in the 2000s. [@popova2018] cited a 2016 study by UN Children's Fund Bulgaria indicating that one-fifth of Bulgarian children had one or both parents working abroad. In China, the 2015 one-percent National Population Sample Survey showed that one out of four children lived under parental absence induced by migration [@nbs2017].

Parental migration impacts child-and-adolescence development by exerting different effects that work oppositely [@vanhook2020]. Consider children's education. Labor migrants send remittances back home, boosting the family income. Higher income can fund children's schooling as well as improve their nutritional and living conditions. All these may enhance children's educational outcome. The inflow of remittances, however, cannot substitute for the loss of parental care. Parental migration halts the day-to-day, face-to-face contact between parents and children. It decreases the amount and quality of parental guidance, protection, and support. All these may hinder children's school performance. It is uncertain how these two opposing effects shape children's educational performance, which merits empirical investigation.

A challenge arises in when we want to assess the causality between parental migration and children's education. Experiments can unveil causal relations, but it is unfeasible and unethical to design an experiment that randomly assigns parents to migrate without their children. Statistical advances in causal inference provides an alternative approach. We can infer causality from a quasi-experimental design -- that is, only if the cause is as good as randomly assigned to individuals in the analysis, conditional on identification assumptions.

For studying left behind children, a suitable case is China. The country has seen a huge scale of internal migration and salient challenges facing migrant parents and left behind children. The National Bureau of Statistics reported that the number of in-country migrant workers exceeded 292.5 million by 2021[@nbs2022]. According to [@gu2021], current policy still limits internal migrants' access to public services and social benefits outside their hometowns, including public school for their children. Facing financial and policy constraints, many migrant laborers have to leave children in their original domiciles. The number of left behind Chinese children reached 68.7 million in 2015 [@nbs2017].

To what extent do parental migration affect the cognitive development and academic performance of left-behind children in the short term? I analyze the data from a nationally representative two-wave panel survey. The sample includes children in both urban and rural areas. I develop a difference-in-differences design for causal inference. The dependent variables include children's cognitive test score and Chinese, mathematics, and English exam scores.

@@

Researchers obtained mixed results regarding how migration-induced parental absence affects children's education. Using data collected from junior high school students, [@li2017] found that parental migration negatively impacts left-behind children's achievement in standardized mathematical test. [@wang2019] also collected data from students of junior high school, but discovered no overall effect of parental migration on the mathematical test grades of children left behind. [@bai2018] gathered data from primary school pupils. They reported that parental migration "does not appear to have a negative effect" on the standardized English test scores of left-behind children, and "in some cases even appears to have a positive effect" (p. 1165). The inconsistency can result from factors such as the left-behind children's age and gender, the migrant parent's original domicile and eventual destination, and the arrangement of which parent migrate for how long.

Between parental migration and children's education, the mechanisms trace to two theoretical frameworks. One model, the generation model, holds that parental migration brings forth greater resources in the form of economic capital. Economic capital is, in Bourdieu's words, "immediately and directly convertible into money and may be institutionalized in the form of property rights" [-@bourdieu2002, p.16]. Economic capital is transferable to human capital, which is the stock of "skills, talents, health, and expertise" in people that enhance their productivity [@goldin2016; @botev2019]. [@chang2019] provided evidence through interviews with migrant workers and their family members. They said that the city offered more jobs and higher pay than did the countryside. Oftentimes, they found that labor migration presented the only way to sustain their children's education and livelihood.

The other theoretical model is the family disruption model. Children flourish under parental nurturing care. This term refers to the "stable environments that promote health and adequate nutrition, protect from threats, and provide opportunities for learning and responsive, emotionally supportive and developmentally enriching relationships" [@black, p.1]. Parental migration decreases nurturing care, thereby hindering children's developmental outcomes. The loss of nurturing care hinders children's accumulation of human capital. [@hong2019] talked to a group of left behind children in grade nine. These students expressed the utter sense of loneliness "with no advocate, no support and with no one looking out for them" (p. 13). Many of them felt indifferent to school grades, and some even longed to ditch school to pursue a free, independent life (p. 14).

I hypothesized that in the short term, parental migration negatively affect the cognitive development and academic performance of children left behind. This is because parental migration immediately disrupts family structure, resulting in lower academic performance of children. It takes longer time for parental migration to generate resources that improve children's performance at school.

Researchers have studied whether the impact differ between mother absence and father absence in the context of labor migration. [@xu2019] adopted a fixed-effects propensity score weighting model on cross-sectional data collected from seventh and ninth graders in rural areas across China. They found that the absence of father only or both parents had "little or no association with negative outcomes" on children's academic and cognitive performances. Only-mother absence, however, showed a "strong association with negative outcomes" (p. 1646). [@chen2019] conducted structural equation modeling on the cross-sectional data collected from fourth to seventh graders in the countryside of one province. They found that mother migration was "negatively associated with children's social competence and academic performance," while father migration had no direct effects on the two outcomes (pp. 860-861).

I hypothesized that in the short term, mother migration negatively affect the cognitive development and academic performance of children left behind, while father migration does not show a effect. This relates to the social norms that designate men as breadwinners and women as caregivers. Studies noted that in some Asian and Latin American societies, children are more likely to accept fathers' migration than mothers' [@murphy2022]. Left behind children may even resent their migrant mothers, viewing them as transgressing on the expected care-giving roles.

The research on China's left behind children has three major shortcomings. First, most literature focuses on children left behind in rural areas while neglecting the urban counterparts. As [@nbs2017] stated, the number of urban left behind children had been growing, and the trend would continue. The second shortcoming lies in data and sampling. Some studies draw from cross-sectional data, making it hard to infer causal relations [@zhou2014; @jin2020]. Others suffer from a limited sample size of less than one thousand participants [@liu2021; @yang2022] or a limited geographic coverage of one or a few towns [@yang2022; @shu2021; @jin2020]. The third shortcoming lies in the selection and measurement of outcome. [@chang2019; @wang2019] measure students' educational performance with the test score of only one academic subject, namely mathematics. The test results of multiple subjects should be included, as parental migration's impact may vary across subject areas.

@@

I obtained the data from the China Education Panel Survey (CEPS). It is a nationally representative, school-based survey featuring junior high school students. The survey administrator is the National Survey Research Center at Renmin University of China. The research team adopted the approach of multi-stage stratified probability proportional to size sampling. The selected participants filled a paper-and-pencil questionnaire. The baseline survey took place in the 2013-2014 academic year with a sample size of about twenty thousand students in seventh and ninth grades. These students were nested in 438 classrooms of 112 schools in 28 urban districts or rural counties. I only kept the 9,449 observations of seventh grade students and their parents, who completed the follow-up survey in the 2014-2015 academic year.

@@

I explored two categories of adolescence development: academic and cognitive abilities. One key dependent variable, the student's cognitive skill, is measured by a fifteen-minute standardized test. The test covered three types of reasoning: verbal, visuospatial, and numerical. The raw scores of cognitive skill were then standardized. The other set of dependent variables concerned students' academic performance. This was measured by the mid-term exam scores in three subjects: Chinese, mathematics, and English. I calculated the standardized exam grades at the school-by-grade level to facilitate comparison across schools and counties. I removed the observations with missing values in any of the test score mentioned above. My choice of standardized exam grades over raw grades follows previous research [@young2018; @xu2019].

The key independent variable, the parental migration status, was identified by two items in the parent survey. In both the baseline and the follow-up surveys, parents specified the family members living in the same household at the time. I constructed six treatment dummy variables measuring migration arrangements: both parents absent, only father absent, only mother absent, father absent (unconditional), mother absent (unconditional). These arrangements can overlap with each other. The control group consisted of parents staying with their children throughout at both the baseline and the follow-up surveys. I then removed the observations that reported other types of parental absence induced by divorce or death.

The control variables at the student level include: student's age, gender, ethnicity, self-rated health, hukou type, and the number of siblings. I obtained the age by subtracting the student's birth year from the year 2013, the start of the baseline academic year. I coded students' gender as male = 0 and female = 1, ethnic minority status as Han Chinese = 0 and other ethnicity = 1, hukou status as rural (agricultural) hukou = 1 and other types = 0. I obtained the number of siblings of students through two items. Students who answered yes to the question "Are you the only child of your family" had no sibling. Those who answered no to this question were instructed to proceed to the next question, "How many full or half siblings do you have" -- elder brother, younger brother, elder sister, and younger sister, respectively. In these four columns, I transformed `NA` to zero. Then I added the four columns as the number of siblings of a student.

The control variables at the household level include the internet assess at home, the self-rated number of books at home, and the years of education of the best-educated parent. These variables can approximate the level of human capital in a household. The internet assess was measured by the item "Do [sic] your family own a computer and have an access to the Internet". Those who answered "Yes, we have both" were coded as 1, and others as 0. The number of books was measured by the item "How many books do your family own? (not including textbooks or magazines)". Students chose from a range of 1 as "very few" to 5 as "a great number". The years of education of the best-educated parent were measured by two items. Students indicated their father's and mother's educational level, respectively. I transformed these two variables into years of education and retained the larger value.

@@

I adopted `r cite_r("r-references.bib")` for all the analyses, the `fixest` package for building the econometric models [@bergé2018], and the `modelsummary` package for presenting the results [@arel-bundock2022]. I built a two-way fixed effect model in a difference-in-differences (DID) design, following Bai and colleagues' [-@bai2018; -@bai2020] approach. The model served to analyze the educational outcome of children newly left behind by parents vis-a-vis that of children staying together with parents. I first specified an unrestricted and unadjusted model: $$\Delta score_{i,s} = \alpha + \beta \cdot migr_{i,s} + \gamma \cdot FE_{s} + \lambda \cdot score_{i,s;base} + \varepsilon_{i,s}$$For student $i$ in school $s$, $\Delta score_{i,s}$ is the change in standardized test score between baseline and follow-up surveys, $migr_{i,s}$ is the treatment dummy variable, $FE_{s}$ is the school-level fixed effect, and $score_{i,s;base}$ is the standardized test score at baseline. The model is unrestricted because it does not restrict on the coefficient associated with the baseline scores. The model is unadjusted as it does not adjust for additional covariates. Theoretically, it is unnecessary to include covariates that vary over group but remain consant over time; They would cancel out in the two-way fixed effect model. The standard errors were clustered at the school level.

In addition, I presented an unrestricted and adjusted DID model: $$\Delta score_{i,s} = \alpha + \beta \cdot migr_{i,s} + \gamma \cdot FE_{s} + \lambda \cdot score_{i,s;base} + \theta \cdot X_{i,s} + \varepsilon_{i,s}$$where $X_{i,s}$ is the vector of covariates capturing the characteristics of children and their households. The control variables were measured in the baseline survey. This version of model lifts the restriction that covariates from the baseline survey would be associated with a coefficient that equals one.

@@

```{r}
table_unres_cog
```

```{r}
table_unres_chn
```

```{r}
table_unres_mat
```

```{r}
table_unres_eng
```

```{r}
table_adj_unres_cog
```

```{r}
table_adj_unres_chn
```

```{r}
table_adj_unres_mat
```

```{r}
table_adj_unres_eng
```

@@

>>>>

---
title             : "Newly Left-Behind Children's Educational Performance Drops Slightly After Parents Migrated"
shorttitle        : "Left-Behind Children"

author: 
  - name          : "Jiancheng Gu"
    affiliation   : "1"
    address       : "De Boelelaan 1105, 1081 HV - Amsterdam, The Netherlands"
    email         : "j4.gu@student.vu.nl"

@@

floatsintext      : yes
linenumbers       : no
draft             : no
mask              : no

@@

```{r setup, include = FALSE}
load('.Rdata')
library("papaja")
library("kableExtra")
r_refs("references.bib")
```

@@

Millions of children are resided in their home communities but separated from their parents as either one or both parents have migrated for work [@antia2020]. These are the so-called "left-behind children" or "stay-behind children." They are common in migrant-sending regions, such as Latin America, Sub-Sahara Africa, East Europe, and large parts of Asia. Researchers estimated that one-sixth of Mexican children [@dewaard2018], one-fifth of Bulgarian children, [@popova2018], and one-fourth of Chinese children were living apart from their migrating parents [@nbs2017].

How are stay-behind children faring when parents have migrated? Scholars have obtained mixed results due to the differential effects of parental migration. Consider children's education. According to [@vanhook2020] and [@debrauw2019], labor migrants send remittances back home, which can fund children's schooling as well as improve their nutritional and living conditions. The inflow of remittances, however, cannot substitute for the loss of parental care. Parental migration decreases the amount and quality of parental guidance, protection, and support. It is uncertain how these two opposing effects together shape children's educational performance, which merits empirical investigation.

A challenge arises in when we want to assess the causality between parental migration and children's education. Experiments can unveil causal relations, but it is unfeasible and unethical to design an experiment that randomly assigns parents to migrate without their children. Alternatively, we can infer causality from a quasi-experimental design -- that is, when the cause is on par with random assignment to participants in the analysis, conditional on identification assumptions [@cunningham2021].

To what extent do parental migration affect the educational performance of left-behind children in the short term? I choose China as the case and analyze data from a nationally representative two-wave panel survey. The sample includes children in both urban and rural areas. I develop a difference-in-differences model for causal inference and apply propensity score matching for robustness check. The outcome was measured with children's cognitive test score and academic exam scores. I demonstrate that parental migration reduces the educational performance of children staying behind, but only to a small extent.

@@

## Theoretical Framework

Between parental migration and children's education, the mechanisms trace to two theoretical frameworks. One is the resource generation model. It holds that parental migration brings forth greater resources in the form of economic capital. Economic capital is, in Bourdieu's words, "immediately and directly convertible into money and may be institutionalized in the form of property rights" [-@bourdieu2002, p.16]. Economic capital is transferable to human capital, which is the stock of "skills, talents, health, and expertise" in people that enhance their productivity [@goldin2016; @botev2019]. [@yang2008] noted that in the Philippes, a sharp depreciation of peso increased the amount of remittances from migrant laborers working abroad. This subsequently boosted children's school enrollment and educational expenditure. In China, [@chang2019] interviewed migrant workers and their family members. Oftentimes, they found that labor migration presented the only way to sustain their children's education and livelihood.

The other theoretical model is the family disruption model. Children flourish under parental nurturing care. This term refers to the "stable environments that promote health and adequate nutrition, protect from threats, and provide opportunities for learning and responsive, emotionally supportive and developmentally enriching relationships" [@black2021, p.1]. Parental migration decreases nurturing care, thereby hindering children's developmental outcomes. The loss of nurturing care slows children's accumulation of human capital. Hong talked to a group of left behind children at a rural school [@hong2019]. These students expressed the utter sense of loneliness "with no advocate, no support and with no one looking out for them" (p. 13). Many of them felt indifferent to school grades, and some even longed to ditch school to pursue a free, independent life (p. 14).

Researchers obtained mixed results regarding how parental migration affects children's education. [@botezat2020] studied the case of Romania and found a positive effect of parents' migration on academic grades of adolesents aged eleven to fifteen. [@wang2019] collected data from junior high school students in central China and found no effect of parental on left-behind children's achievement in mathematical test. Studying four Chinese provinces, [@debrauw2017] showed that as ample migration opportunities appeared in villages, enrollment to senior high school decreased. [@nguyen2016] demonstrated that parental migration was associated with lower cognitive capabilities among Indian and Vietnamese children aged five to eight. The inconsistency can result from factors such as the left-behind children's age and gender, the migrant parent's original domicile and eventual destination, and the arrangement of which parent migrate for how long.

Parental migration immediately disrupts family structure, which in turn lowers academic performance of children. It takes longer time for parental migration to generate resources that improve children's performance at school. I propose the first hypothesis:

$H_1$ In the short term, parental migration negatively affect the cognitive development and academic performance of children left behind.

Researchers have studied whether the impact differ between mother absence and father absence in the context of labor migration. Xu and others [-@xu2019] adopted a fixed-effects propensity score weighting model on cross-sectional data collected from seventh and ninth graders in rural areas across China. They found that the absence of father only or both parents had "little or no association with negative outcomes" on children's academic and cognitive performances. Only-mother absence, however, showed a "strong association with negative outcomes" (p. 1646). Chen and others [-@chen2019] conducted structural equation modeling on the cross-sectional data collected from fourth to seventh graders in the countryside of one province. They found that mother migration was "negatively associated with children's social competence and academic performance," while father migration had no direct effects on the two outcomes (pp. 860-861).

The social norms designate men as breadwinners and women as caregivers. Studies noted that in some Asian and Latin American societies, children are more likely to accept fathers' migration than mothers' [@murphy2022]. Left behind children may even resent their migrant mothers, viewing them as transgressing on the expected care-giving roles. I propose the second hypothesis:

$H_2$ In the short term, mother migration negatively affect the cognitive development and academic performance of children left behind, while father migration does not show a effect.

## China's Context

Three features make China a suitable case for studying left behind children. The first feature is the huge scale of internal labor migration. According to the National Bureau of Statistics, the year 2021 recorded 292.5 million in-country migrant workers [@nbs2022]. The number of left behind Chinese children reached 68.7 million in 2015 [@nbs2017]. The second feature is the neoliberal welfare regime. Neoliberalism extends the logic of the market to most spheres of human life. Driven by the neoliberal doctrine, the government shifts its responsibilities of social services and welfare to individuals and communities. Since China initiated pro-market reforms in the 1980s, the state has reduced its welfare provisions and passed them to individuals and families [@duckett2020]. The third is the authoritarian control against citizens. According to Gu [-@gu2021] and Zhang [-@zhang2018], the hukou (household registration) system controls people's mobility by binding a household's public welfare access with its place of registry. One can move out of town freely but cannot change hukou freely. For a rural household member moving to the city, it takes much time and effort to apply for urban hukou status. Without an urban hukou they cannot send children to public school in the host city. Facing financial and policy constraints, many migrant laborers have to leave children in their original domiciles.

The research on China's left behind children has three major shortcomings. First, most literature focuses on children left behind in rural areas while neglecting the urban counterparts. [@nbs2017] stated that the number of urban left behind children had been growing and expected that the trend would continue. The second shortcoming lies in data and sampling. Some studies draw from cross-sectional data, making it hard to infer causal relations [@zhou2014; @jin2020]. Others suffer from a limited sample size of less than one thousand participants or a limited geographic coverage of one or a few towns [@liu2021; @yang2022; @shu2021; @jin2020]. The third shortcoming lies in the selection and measurement of outcome. [@chang2019; @wang2019] measured students' educational performance with the test score of only one academic subject, namely mathematics. The test results of multiple subjects should be included to measure comprehensive academic achievement.

@@

I obtained the data from the China Education Panel Survey (CEPS) [@nsrc]. It is a nationally representative, school-based survey featuring junior high school students. The survey administrator is the National Survey Research Center at Renmin University of China. The research team adopted multi-stage stratified probability proportional to size sampling and administered a paper-and-pencil questionnaire to each participant. The baseline survey took place in the 2013-2014 academic year with a sample size of about twenty thousand students in the seventh and ninth grades. These students were nested in 438 classrooms of 112 schools in 28 urban districts or rural counties. The follow-up survey in the 2014-2015 academic year lost track of the ninth grade cohort and only surveyed `r printnum(nrow(d))` of the seventh grade cohort. I further restricted the sample to students staying with both parents at the original domicile at the baseline survey. To deal with missing values, I adopted list-wise deletion for dependent, independent, and control variables. The final sample size amounted to `r printnum(nrow(d_cl))` .

@@

I explored two categories of educational performance: cognitive and academic abilities. One key dependent variable, the student's cognitive skill, was measured by a fifteen-minute standardized test. The test covered three types of reasoning: verbal, visuospatial, and numerical. The CEPS research team calculated the raw score based on the Item Response Theory and then standardized the score with a mean of zero and standard deviation of one.

Another dependent variables concerned students' academic performance. This was measured by the total scores of mid-term exams in Chinese, mathematics, and English as provided by each school. These three subjects constitute the core parts in the senior high school entrance exam. Junior high school students study these subjects throughout three academic years. Within each cohort in one particular school, teachers adopt the same syllabus and administer the same exams during a given assessment period. Therefore, test scores for core subjects are reliable indicators of learning performance in the same school and cohort. I aligned different grading scales (full mark of 100, 120, and 150 for one subject) to the hundred-point scale. After the conversion, I checked for abnormal values outside the 0-100 range. `r printnum(num.outlier$n)` observations are outliers, accounting for less than one percent of the sample. I trim the values of subject exam score with minimum of zero and maximum of one hundred. In this way, I can limit the impact of outliers while retaining these observations in the sample. After trimming, I aggregated the subject exam scores to obtain a total score of a 300-point scale.

The key independent variable, the parental migration status, was identified by two items in the parent survey. In both the baseline and the follow-up surveys, parents specified the family members living in the same household at the time. I constructed four treatment dummy variables measuring migration arrangements: any parents absent, only father absent, only mother absent, and both parents absent. These arrangements are mutually exclusive. The control group consisted of parents staying with their children throughout at both the baseline and the follow-up surveys. I then removed the observations that reported parental divorce or death. These type of parental absence are irrelevant to migration.

@@

I adopted `r cite_r` for all the analyses, the `fixest` package for building the econometric models [@bergé2018], and the `modelsummary` package for presenting the results [@arel-bundock2022]. I built a two-way fixed effect model in a difference-in-differences (DID) design, following Bai and colleagues' [-@bai2018; -@bai2020] approach. The model served to analyze the educational outcome of children newly left behind by parents vis-a-vis that of children staying together with parents. I first specified an unrestricted and unadjusted model: $$\Delta score_{i,j,k} = \alpha + \beta \cdot migr_{i} + \gamma \cdot C_{i} + \lambda \cdot C_{j} + \theta \cdot score_{i,j;1} + \varepsilon_{i,j,k}$$ For student $i$ in class $j$ in baseline and class $k$ in endline, $\Delta score_{i,j}$ is the change in test score between baseline and follow-up surveys, $migr_{i}$ is the treatment dummy variable, $C_{j}$ and $C_{k}$ are respectively the classroom fixed-effects at baseline and follow-up, and $score_{i,j,k;1}$ is the test score at baseline. The model is unrestricted because it does not restrict on the coefficient associated with the baseline scores. The model is unadjusted as it does not adjust for additional covariates. Theoretically, it is unnecessary to include covariates that vary over group but remain constant over time; They would cancel out in the two-way fixed effect model [@huntington-klein]. The standard errors were clustered at the classroom level.

In addition, I present an unrestricted and adjusted DID model: $$\Delta score_{i,j,k} = \alpha + \beta \cdot migr_{i} + \gamma \cdot C_{i} + \lambda \cdot C_{j} + \theta \cdot score_{i,j,k;1} + \zeta \cdot X_{i} + \varepsilon_{i,j,k}$$ where $X_{i}$ is the vector of covariates capturing the characteristics of children, their parents, and their households. The control variables were measured in the baseline survey. This version of model lifts the restriction that covariates from the baseline survey would be associated with a coefficient that equals one.

@@

## Descriptive Statistics

The table below summarizes the changes in migration status in the sample. All `r printnum(nrow(d_cl))` children were living together with parents in one household in baseline, but about fifteen percent saw at least one parent migrated during the two waves of survey. In these new migrant households, only-mother migration and only-father migration each accounted for two out of five cases. The remaining one-fifth was both-parents migration.

```{r}
table_independent_variable
```

: Parental migration status at follow-up survey.

The table below provides the summary statistics.

```{r}
table_dependent_variable
```

: The independent variables

```{r}
table_control_variable
```

: The control variables

## Regression Analysis

The unrestricted DID models support my hypothesis one -- newly left behind adolescents are more likely to perform worse in cognitive and academic abilities than their non left behind counterparts. If a household sees any parent migrated, the child's cognitive test score and total exam grade would decrease relative to that of children whose parents stay with them, holding everything else constant. Parental migration is linked to is `r round(coef(did_cog)[1],2)` SD in cognitive test score and `r round(coef(did_cog)[1],2)` points in academic exam grades. Both relationships are slight in strength.

The results only partially support the hypothesis two -- mother migration has a more negative effect than father migration. This holds true for academic abilities. Children from only mother migrate households are associated with `r round(coef(did_cog_m_only)[1],2)` points relative to that of children whose parents stay along. The performances of children with both parents migrated or only father migrated do not show statistically significant difference from the non left behind children. Regarding cognitive abilities, the reverse is true. Children with only mother migrated do not show statistically significant difference from the non left behind children. Children with only father absent are associated with (`r round(coef(did_cog_f_only)[1],2)` SD), and both parents absent (`r round(coef(did_cog_both)[1],2)` SD) compared to the non left behind children.

```{r}
table_cog
```

```{r}
table_ttl
```

The adjusted DID models produce similar results compared to the unadjusted models, with only one exception. That is, for the cognitive skills, the both-parent-absent coefficient saw a change of statistical significance level from five percent level to ten percent level.

Some predictors are strong in both models of cognitive and academic abilities. These are children's age and educational asiration as well as parents' expectation on children. In terms of age, the older students' scores drop relatively more than younger students'. One year older in age is associated with `r round(coef(did_adj_cog)[3],2)` SD for standardized cognitive test and `r round(coef(did_adj_ttl)[3],2)` points for academic exams. Parent's expectation for children comleting university and children's own aspiration for finishing university are associated with higher scores. The effect sizes are `r round(coef(did_adj_cog)[15],2)` SD and `r round(coef(did_adj_cog)[12],2)` SD for standardized cognitive test, respectively; and `r round(coef(did_adj_ttl)[15],2)` points and `r round(coef(did_adj_ttl)[12],2)` points for academic exams, respectively.

Several predictors are strong in the model of academic grades but not in that of cognitive scores. These are children's gender and self-reported health, and the years of parental education. Girls outperform boys by `r round(coef(did_adj_ttl)[4],2)` points in academic exams, but do not show advantage in cognitive test. Students who consider themselves healthier are associated with higher scores in academic exams. Parents with more education are associated with their children's higher scores in academic exams. If the years of parental education of the best-educated among the two increase by one, it would be linked to a `r round(coef(did_adj_ttl)[14],2)` points increase in the child's academic exam grade. These three predictors are not significant for cognitive outcomes.

Preschool attendance is associated with higher cognitive outcomes `r round(coef(did_adj_cog)[8],2)` SD, but its effect for academic outcomes is not significant.

```{r}
table_adj_cog
```

```{r}
table_adj_ttl
```

## Matching

Researchers noted that migrants represent a positively selected group compared to non-migrants [@vanhook2020]. One technique to tackle the selectivity issue is matching. This method works by pairing a unit in the treatment group with one or several units in the control group based on observable covariates. Popular choices include coarsened exact matching (CEM) [@waldman2022], nearest-neighbor propensity score matching (PSM) [@xu2015], and nearest-neighbor Mahalanobis distance matching (MDM) [@guo2021]. CEM is superior in achieving balance as it optimizes the entire joint distribution of covariates. The downside is that it leaves many units without an exact match and discards these unmatched units. The nearest-neighbor matching examines all the treated units and selects the closest eligible control unit to be paired. The eligibility as the "closest" can be determined with the differences in Mahalanobis distance or in propensity score. Propensity score denotes the probability of receiving treatment based on the specified covariates. Nearest-neighbor PSM does not try to optimize any condition as each pairing ignores how other units will be or have been paired. This matching method may result in more "imbalance, model dependence, and bias" on highly balanced data [@king2019].

I conducted matching with all the covariates used in the adjusted DID model. I first attempted CEM, which left most units unmatched and resulted in a tiny sample. Then I combined exact matching (based on the student's county of residence) and nearest neighbor PSM (based on the other covariates), which is the configuration of Bai and others [-@bai2018]. Exact matching based on county of residence preserves the original geographic distribution. The propensity score was estimated using a logistic regression of the treatment on the covariates. The matching ratio was set at one treatment unit with up to three control units.

Alike are the results with matching and those without. In the models of cognitive abilities, I obtained negative, statistically significant coefficients on the treatment variables in the any parent migrated household and only father migrated households. In the models of academic abilities, the treatment variables' coefficients are negative and statistically significant for the any parent migrated households and only mother migrated households. The magnitudes of the coefficients are also similar.

```{r}
table_m_cog
```

```{r}
table_m_ttl
```

```{r}
table_m_adj_cog
```

```{r}
table_m_adj_ttl
```

@@

This thesis proposed a model to infer causality between parents' migration and their stay-behind children's educational performance in the short run. I built econometric models in the difference-in-differences design to analyze data from the 2013-2014 and 2014-2015 rounds of China Education Panel Survey. Results suggest that parental migration is, in the short run, negatively associated with children's educational performance as measured by academic exams and standardized cognitive tests. I further examined whether the effect varies across different arrangements of parental migration. Results indicate that father-only migration is negatively associated with children's cognitive abilities, while mother-only migration is negatively associated with children's academic abilities. In all scenarios, the educational performance of newly left behind children only decreased slightly compared to that of non-left-behind peers.

My research has several limitations in sample selection. The CEPS questionnaire asks if parents live in the household, but does not ask why parents are absent. Given such constraint, I can only identify parental migration through such a condition: parental absence from a household without parental divorce or death. It is possible that some parents are jailed or hospitalized. These types of absence would be wrongly attributed to migration. The CEPS also lacks information on parents' migration history. Possibly, some parents out-migrated at an earlier date, returned before the baseline survey, and migrated away again during the survey period. Their children might better cope with parental absence during the survey period, yet they would be mistaken as newly left behind.
